---
title: "Assignment 1"
author: "Philip"
date: "03 October 2016"
output: html_document
---


## Extract and Loading the Data
This section creates a table which further analysis will be perfromed, /activity/.
First appropriate extracts the activity file contained in the GitHub repository downloaded on *Insert Date here*.

``` {r libraries, echo=TRUE, results="hide"}
library(lubridate)
```


``` {r download_repo_zip, echo=TRUE, results="hide"}
## R code to download file from R D Peng's repo and unzips it
unzip("activity.zip")

```

The data was downloaded and read in as a dataframe. The data field was converted to a date type for further grouping.
```{r read_data, echo=TRUE, results="hide"}
activity<-read.csv(file="activity.csv",stringsAsFactors = FALSE)
activity$date<-ymd(activity$date)
summary(activity)
```


## Exploring the Data
For the following analysis missing values were ignored. It is additionally assumed
that days with 0 steps is a result of missing data. NAs were removed to prevent
operating on NAs should the sensors report NAs for any particular interval in a day.
```{r initialExplore, echo=FALSE}
AverageStepspDay <- tapply( 
                            activity$steps
                            ,activity$date
                            ,function (x) sum(x, na.rm = TRUE)) ## Removed NAs
```


Mean and Median were determined ignoring days with 0 steps.
```{r histogram1}
StepHist<-hist(AverageStepspDay 
                ,main="Distribution of total number of steps per day"
                ,xlab="Total number of steps in a day")

## Account for 0s induced by likely missing data
AdjustedFrequency <- StepHist$counts[1]-
                        length(x = AverageStepspDay[AverageStepspDay==0])
                        ## Determine Frequency without 0s
segments(StepHist$breaks[1],AdjustedFrequency,
         StepHist$breaks[2],AdjustedFrequency)
text(x = StepHist$mids[1],y = (AdjustedFrequency+StepHist$counts[1])/2,
     labels="0 Steps")

meanSteps <- mean(AverageStepspDay[AverageStepspDay!=0])
medSteps <- median(AverageStepspDay[AverageStepspDay!=0])

abline(v = c(meanSteps,medSteps),col=c("blue","red"),lty=c(1,2))
text(x=meanSteps,y = mean(StepHist$counts)+1 #Quick Adjustment on text position
     ,pos=2
     ,labels = paste("Mean",round(meanSteps,digits = 0))
     ,col = "blue")
text(x=medSteps,y = mean(StepHist$counts)+1 #Quick Adjustment on text position
     ,pos=4
     ,labels = paste("Median",round(medSteps,digits = 0))
     ,col = "red")

```

The histogram clearly details the 

